Measurement of ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross sections in the LArIAT experiment

### 2018-05-30 Chapter-5/5-Preparatory.tex
s

\section{Cross Section Analyses Data Sets}\label{sec:dataSet}
We choose LArIAT Run-II as the data period for the  ($\pi^{-}$,Ar) and (K$^{+}$,Ar) total hadronic cross section analyses. 
Data taking for the this period started on 03/15/2016  and ended on 07/31/2016. 
Since we are interested in beamline and TPC information, we ask basic requirements on the operational status of the time of fight, wire chambers and TPC to form the good run list for this period, which we informally call ``lovely runs".

The subset of lovely runs  chosen for the  ($\pi^{-}$,Ar) total hadronic cross section analysis includes only the -60A and -100A magnet configurations in negative polarity, even if LArIAT explored several other beamline configurations during Run-II. The -60A and -100A combined data set accounts for approximately 90\% of the total Run-II negative polarity runs.   The  choice of the main two beamline settings limits the need for the production of many different MC sets and related corrections, still maintaining a high number of events. 
%Since the production of beamline Monte Carlo depends on the wanted beamline configuration, the choice of only two beamline settings limits the need for beamline MC production.

@@

For the first measurements in LArIAT that uses both beamline and TPC information, we choose strict requirements on the reconstruction of the WC tracks, the so-called ``Picky Track" sample (see Section \ref{sec:MWPCfunc}). This choice presents two advantages:  the uncertainty on the momentum reconstruction for the ``Picky Tracks" sample is smaller compared to the ``High Yield" sample, and the comparison with the beamline MC results is straightforward. A possible future update and cross check of these analysis would be the use of the High Yield sample, where the statistics is about three times higher.

@@

\section{Construction of a Monte Carlo Simulation for LArIAT}\label{sec:MCSet}
For the simulation of LArIAT events and for the simulation of the datasets' particle make up, we use a combination of two MC generators: the G4Beamline Monte Carlo and the Data Driven single particle Monte Carlo (DDMC). We use the G4Beamline MC to simulate the particle transportation in the beamline and calculate the particle composition of the beam just after the fourth Wire Chamber (WC4). In order to simulate the beamline particles after WC4 and in the TPC, we use the DDMC.

\subsection{G4Beamline}\label{ch:beamlineComposition}
G4Beamline simulates the beam collision with the LArIAT secondary target, the energy deposited by the particles in the LArIAT beamline detectors, and the action of the LArIAT magnets, effectively accounting for particle transportation through the beamline from the LArIAT target until ``Big Disk", a fictional, void detector located just before the LArIAT cryostat. 
 At the moment of this writing, G4Beamline does not simulated the responses of the beamline detectors. It is possible to interrogate the truth level information of the simulated particles in several points of the geometry. In order to ease the handshake between G4Beamline and the DDMC, we ask for the beam composition just after WC4.
Since LArIAT data are taken under different beam conditions, we need to simulate separately the beam composition according to the magnets' settings and the secondary beam intensity with G4Beamline. For the pion cross section analysis the relevant beam conditions are  secondary beam energy of 64 GeV, negative polarity magnet with current of 100 A and 60 A. For the kaon cross section analysis the relevant beam conditions is a secondary beam energy of 64 GeV, positive polarity magnet with current of 100 A.

\subsubsection{Beam Composition for Negative Pion Cross Section}
Even if pions are by far the biggest beam component in negative polarity runs, the LArIAT tertiary beam is not a pure pion beam. While useful to discriminate between pions, kaons, and protons, the beamline detectors are not sensitive enough to  discriminate among the lighter particles in the beam: electrons, muons and pions fall under the same mass hypothesis. Thus, we need to assess the contamination from beamline particles other than pions in the event selections used for the pion cross section analysis and correct for its effects. The first step of this process is assessing the percentage of electrons and muons in the $\pi/\mu/e$ beamline candidates via the G4Beamline MC. 
Since the beamline composition is a function of the magnet settings, we simulate separately events for magnet current of -60A and -100A. 
Figure \ref{fig:BeamComposition} shows the momentum predictions from G4Beamline overlaid with data for the 60A runs (left) and for the 100A runs (right). The predictions for electrons, muons and pions have been staggered and their sum is area normalized to data. Albeit not perfect, these plots show a reasonable agreement between the momentum shapes in data and MC.  We attribute the difference in shape to a two approximations performed in the MC. Firstly, G4Beamline lacks the simulation of the WC efficiency which is momentum dependent and leads to enhance the number events in the center of the momentum distribution. Secondly, G4Beamline stop tracking pions and their products if they decay in after WC1; in data, pion decays in flight can still create a tigger if the produced muon travels thought the beamline detectors. In the pion cross section analysis, these differences between data and G4Beamline are accounted for as a systematic uncertainty related to the beam composition (see Section \ref{ch:BKGsubXSPuppa}).
%We attribute  the difference in shape to the lack of simulation of the WC efficiency in the MC which is momentum dependent and leads to enhance the number events in the center of the momentum distribution.

@@

\subsection{Data Driven MC}\label{sec:DDMC}
The Data Driven single particle Monte Carlo (DDMC) is a single particle gun which simulates the particle transportation from WC4 into the TPC leveraging on the beamline data information. The DDMC uses the data momentum and position at WC4 to derive the event generation: a general sketch of the DDMC workflow is shown in Figure \ref{fig:DDMCSketch}.

When producing a DDMC sample, beamline data from a particular running period and/or running condition are selected first. For example, data for the negative 60A runs and for the negative 100A runs inform the event generation stage of two different DDMC samples. Figure \ref{fig:DDMCQuantities}  schematically shows the data quantities of interest leveraged from data: the momentum ($P_x, P_y, P_z$) and position ($X, Y$) at WC4. For each data event, we obtain the  particle position ($X, Y$) at WC4 directly from the data measurement; we calculate the components of the momentum using the beamline measurement of the momentum magnitude in conjunction with the hits on WC3 and WC4 to determine the direction of the momentum vector, as described in section \ref{sec:MWPCfunc}. The momentum and position of the selected data form a 5-dimensional tupla, which we sample thousands of times through a 5-dimensional hit-or-miss sampling procedure to generate the MC events. This sampling generates MC events  with the same momentum and position distributions as data, with the additional benefit of accounting for the correlations between the $P_x, P_y, P_z, X, Y$ variables.  As an example, the results of the DDMC generation compared to data for the kaon +100A sample are shown in figure \ref{fig:DDMCComparison} for the $P_z$, $X$ and $Y$ distributions; as expected, MC and data agree within the statistical uncertainty by construction. A LArSoft simulation module then launches single particle MC from z = -100 cm (the location of the WC4) using the generated events. The particles are free to decay and interact in their path from WC4 to the TPC according to the Geant4 simulation.

Using the DDMC technique ensures that the MC and data particles have very similar momentum, position and angular distributions at WC4 and allows us to use the MC sample in several occasions: to estimate the background contamination to the pion cross section (see Section \ref{ch:PionXSBkgSub}), to calibrate the energy loss upstream of the TPC (see Section \ref{ch:eloss}),  or to study the tracking and the calorimetric performance (sections \ref{sec:TrackingStudies} and \ref{ch:energyCal}). A small caveat is in order here: the DDMC is a single particle Monte Carlo, which means that the beam pile-up is not simulated.

@@

We use the beamline simulation and the DDMC simulation to estimate the background in the total hadronic pion cross section. Two categories of background exists for the negative pion cross section measurement: the one related to the pion interaction in the chamber, discussed in Section \ref{ch:CaptureAndDecay} and the one related to the beamline contamination, discussed in Section \ref{ch:PionXSBkgSub2}.

\subsection{Background from Pion Capture and Decay}\label{ch:CaptureAndDecay}
Our goal is to measure the total hadronic cross section for negative pions in argon. Since pion capture can be classified as an electromagnetic process and pion decay is a week process,  capture and decay represent unwanted interactions. We present here a study of capture and decay in Monte Carlo and the solution we adopted to mitigate their occurrence in the data sample.

For this MC study, we use a sample of  MC pions generated according to the  $-6$0A beam profile with the DDMC (see Section \ref{sec:DDMC}). It is important to notice that capture occurs predominantly at rest, while decay may occur both in flight and at rest. Thus, we can highly mitigate capture and decay at rest by removing pions which would release all their energy in the TPC and stop. This translates into a momentum selection, where we keep only events whose WC momentum is above a certain threshold. 
Figure \ref{fig:CaptureMom} shows the true momentum distribution for the primary pions\footnote{We use here the Geant4 denomination ``primary" to indicate that the pion considered does not undergo interactions modifying its energy before getting to the TPC. In fact, not every pion shot from wire chamber four will arrive to the TPC as primary,  some will decay or interact before the TPC.}  that arrive to the TPC (pink), that capture (green) or decay (blue) inside the TPC, on a linear scale (left) and on a log scale (right) vertical axis.

In order to choose the selection value for the wire chamber momentum, it is beneficial to estimate the ratio of events which capture or decay that survive the selection in MC as a function of the momentum threshold, and compare it with the survival ratio for all the 60A events. This is done in figure \ref{fig:survRatio}. We define the survival ratio simply  as the number of events surviving the true momentum selection divided by the number of events of that category. We calculate the survival ratio separately for the three event categories explained above: total (pink), capture (green) and decay (blue).
Selecting pions with momentum greater than 420 MeV/c reduces the capture events by ~99\% while maintaining about 80\% of the 60A data sample and almost the entire 100A sample. 
Figure \ref{fig:evtRatio} shows the ratio of events which end their life in capture (green) or decay (blue) over the total number of events as a as a function of the true momentum at wire chamber four. This ratio is slightly dependent on the inelastic cross section implemented in Geant4, as we are able to register a pion capture (or decay) only if it did not interact inelastically in the TPC. We choose a momentum threshold of 420~MeV/c because the percentage of capture events drops below 1\% and the percentage of decays is never above 2\% for momenta greater than 420~MeV/c. After the momentum selection, we evaluate the contribution of capture and decay to be a negligibly small background to the cross section measurement compared to the background related to the beamline which we will address in the next section.

\begin{figure}[]
\centering
\includegraphics[width=15cm]{Chapter-7/Images/CDAsMomentumFunct.png}
\caption{True momentum distribution at wire chamber 4 for every simulated pion arriving in the TPC (pink), ending its life in capture (green) or in decay (blue) in the TPC, linear vertical axis on the left, logarithmic on the right. }
\label{fig:CaptureMom}
\end{figure}

\begin{figure}[]
\centering
\begin{minipage}[t]{0.45\textwidth}
\centering
\includegraphics[width=7.5cm]{Chapter-7/Images/CDThreshold.png}
\caption{Survival ratio as a function of selection threshold on true momentum at wire chamber four for for every simulated pion arriving in the TPC (pink), capture (green) or in decay (blue).   }
\label{fig:survRatio}
\end{minipage}\hfill
\begin{minipage}[t]{0.45\textwidth}
\centering
\includegraphics[width=7.5cm]{Chapter-7/Images/CDRatio.png}
\caption{Ratio between the capture (green) and decay (blue) events over the total number of events as a as a function of the true momentum at wire chamber four.}
\label{fig:evtRatio}
\end{minipage}
\end{figure}

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Contributions from the Beamline Background}\label{ch:PionXSBkgSub2}
We define beamline background every TPC track matched to the WC track which is not a primary pion. Potentially, there are 4 different types of beamline background:
\begin{itemize}
\item[]1) electrons,
\item[]2) muons,
\item[]3) secondaries from pion events,
\item[]4) matched pile up events.
\end{itemize}

The first step to quantify the effect of the beamline background on the pion cross section is to estimate what percentage of events used in the cross section calculation is not a primary pion.  We start by noting that the last type of background, the ``matched pile up" events, is a negligible fraction, because of the definition of the WC2TPC match: we deem the probability of a single match with a halo particle in the absence of a beamline particle\footnote{ Events with multiple WC2TPC matches are always rejected.} negligibly small. %\textcolor{red}{SHOW VTX distribution in WC2TPC match}
As shown in Section \ref{ch:beamlineComposition}, we use G4Beamline to estimate the percentage of pions, muons and electrons at WC4, obtaining the composition shown in Table \ref{tab:beamline}. The next step is to simulate those pions, muons and electrons from WC4 to the TPC with the DDMC and evaluate their contribution to the cross section. To do so, we start by simulating the same number of electrons, muons and pions with the DDMC and we apply the same selection chain (i.e. track multiplicity rejection, WC2TPC acceptance and shower rejection) on the three samples. The number of events per particle species surviving this selection is shown on table \ref{tab:MCafterCutContaminants}. In order to reproduce the closest make up of the beam to data, we weight each event of a given particle species according to the estimated beam composition. In case of 60A runs, for example, the weights are 0.688 for pions,  0.046 for muons  and 0.266 for electrons.

It should be noted that pions may  interact hadronically in the steel or in the non-instrumented argon upstream to the TPC front face while travelling the length of between WC4 and the TPC. Or, they could decay in flight between WC4 and the TPC. One of the interaction products can leak into the TPC and be matched with the WC track, contributing to the pool of events used for the cross section calculation. We call this type of particles ``secondaries" from pion events, with a terminology inspired by Geant4.  We estimate the number of secondaries using the DDMC pion sample.  The percentage of secondaries is given by the number of matched WC2TPC tracks whose corresponding particle is not flagged as primary by Geant4.  The secondary to pion ratio is 4.9\% in the 60A sample and 4.3\% in the 100A sample.

@@

                                                  & MC $\pi^-$   & MC  $ \mu^-$ & MC  $e^-$ & MC  $\pi^-$ & MC  $\mu^-$ & MC  $e^-$  \\
\hline
&  &  &  & & &\\  
Total Initial events                     & 334500  & 334500 & 334500 &344500 &344500& 344500 \\
After Multiplicity Rejection        & 330668  & 333420 & 198065 &326576 &344208& 201380 \\
After WC2TPC Selection          & 218239  & 296333 & 91139  &230418 &300228& 98834  \\
Evts After Shower Rejection     & 208063  & 288914 &  20293 &219882 &293585& 17780  \\
&  &  &  & & &\\  
  \hline
&  &  &  & & &\\  
Selection Survival Rate           &62.3\% & 86.6\% & 6.1\% & 63.8\%& 85.5\%& 5.2\%\\
Beam Composition  @WC4      &  68.8\%   &  4.6 \%  & 26.6 \%    & 87.4 \% & 3.7 \%  & 8.9 \% \\ %
Beam Composition  @TPC FF &  88.5\%   & 8.2\%   & 3.3 \%   & 94.0\%	& 5.3\% & 0.7\%\\
                                                  &                      &                       &                   &                       &                        &\\  
\hline
\end{tabular}
\caption{MC selection flow per particle species.}
\label{tab:MCafterCutContaminants}
\end{table}

@@

We evaluate the beamline background contribution to the cross section by producing the interacting and incident histograms for the events surviving the selection, staggering the contributions for each particle species, as shown in Figure  \ref{fig:stag60A}. From those histograms, we are able to evaluate the contribution of  pions and  beamline backgrounds to each bin of the interacting and incident histograms separately and obtain the relative pion content. The relative pion content in each bin for the interacting and incident histograms represents the correction applied to data. We take here the interacting histogram as example, noting that the derivation of the correction for the incident histogram is identical. The number of entries in each bin of the interacting plot (Figure \ref{fig:stag60A} left) is  $N^{\text{TOT}}_{\text{Int}} (E_{i})$, equal to the sum of the pions and beamline backgrounds in that bin, namely

@@

The pion content is evaluated separately in the interacting and incident histograms. Their ratio determines a correction to the measured raw cross section. 
For example, the measured raw  cross section of a sample with enhanced muons content will tend to be lower than the raw cross section of a muon free sample. This is because most of the muons will cross the TPC without stopping, thus contributing almost exclusively to the incident histogram, forcing the pion content to be lower in the incident histogram than in the interacting; thus, the correction will tend to enhance the cross section.

@@

\section{Tracking Studies}\label{sec:TrackingStudies}
The tracking of hadrons in the TPC determines both the beamline to TPC handshake and the identification of the interaction point within the TPC. Thus, it plays a fundamental role in the cross section measurements. We performed several studies geared towards the optimization of the package for tracking in the TPC. In particular, we studied a suitable set of parameters for the WC2TPC match and we optimized the clustering algorithm to maximize the efficiency of finding the interaction point on MC. Given the technical nature of these studies, we report them in Appendix \ref{ch:AppendixTrack}. 
We report here the evaluation of  the angular resolution of the tracking algorithm in data and MC, due to its implication on the physics measurement.

\subsection{Angular Resolution}\label{sec:angleRes}
Scope of this study is to understand and compare the tracking performances and angular resolution of the TPC tracking on data and MC. 
We use the angular resolution of the tracking to determine  the value of smallest angle that we can reconstruct with a non-zero efficiency, effectively determining a selection on the angular distribution of the cross section measurement due to the tracking performance. %\textcolor{red}{This study is performed on the pion sample, but its results are extrapolated to the kaon case.}

>>>>

\section{Cross Section Analyses Data Sets}\label{sec:dataSet}
We choose LArIAT Run-II as the data period for the  ($\pi^{-}$,Ar) and (K$^{+}$,Ar) total hadronic cross section analyses. 
Data taking for the this period started on 03/15/2016  and ended on 07/31/2016. 
Since we are interested in beamline and TPC information, we ask basic requirements on the operational status of the time of flight counters, wire chambers and TPC to form the good run list for this period, which we informally call ``lovely runs".

The subset of lovely runs  chosen for the  ($\pi^{-}$,Ar) total hadronic cross section analysis includes only the -60A and -100A magnet configurations in negative polarity, even if LArIAT explored several other beamline configurations during Run-II. The -60A and -100A combined data set accounts for approximately 90\% of the total Run-II negative polarity runs.   The  choice of these two main beamline settings limits the need for the production of many different MC sets and related corrections, still maintaining a high number of events. 
%Since the production of beamline Monte Carlo depends on the wanted beamline configuration, the choice of only two beamline settings limits the need for beamline MC production.

@@

For these first measurements that make use of both the LArIAT beamline and TPC information, we choose strict requirements on the reconstruction of the WC tracks, the so-called ``Picky Track" sample (see Section \ref{sec:MWPCfunc}), where we require a single hit in each and every wire chamber detector to reconstruct the WC track. This choice presents two advantages:  the uncertainty on the momentum reconstruction for the ``Picky Tracks" sample is smaller compared to the ``High Yield" sample, and the comparison with the beamline MC results is straightforward. A possible future update and cross check of these analysis would be the use of the High Yield sample, where the statistics is about three times higher.

@@

\section{Construction of a Monte Carlo Simulation for LArIAT}\label{sec:MCSet}
For the simulation of LArIAT events and for the simulation of the datasets' particle make up, we use a combination of two MC generators: the G4Beamline Monte Carlo and the Data Driven single particle Monte Carlo (DDMC). We use the G4Beamline MC to simulate the particle transport in the LArIAT tertiary beamline and calculate the particle composition of the beam just after the last Wire Chamber (WC4). We use the DDMC  to simulate the particles after WC4 along the beamline, close to the beam window in the LAr cryostat and in the TPC.

\subsection{G4Beamline}\label{ch:beamlineComposition}
G4Beamline simulates the beam collision at the LArIAT secondary target, the energy deposited by the particles in the LArIAT beamline detectors, and the action of the LArIAT magnets, effectively accounting for particle transport through the beamline from the LArIAT target until ``Big Disk", a fictional, void detector located just before the LArIAT cryostat. 
 At the moment of this writing, G4Beamline does not simulated the responses of the beamline detectors. It is possible to interrogate the truth level information of the simulated particles in several points of the geometry. In order to ease the handshake between G4Beamline and the DDMC, we ask for the beam composition just after WC4.
Since LArIAT data are taken under different beam conditions, we need to simulate separately the beam composition according to the magnets' settings and the secondary pion beam intensity with G4Beamline. For the pion cross section analysis the relevant beam conditions are  secondary pion beam energy of 64 GeV, negative polarity magnet with current of 100 A and 60 A. For the kaon cross section analysis the relevant beam conditions is a secondary  pion beam energy of 64 GeV, positive polarity magnet with current of 100 A.

\subsubsection{Beam Composition for Negative Pion Cross Section}
Even if pions are by far the biggest beam component in negative polarity runs, the LArIAT tertiary beam is not a pure pion beam. While useful to discriminate between pions, kaons, and (anti)protons, the beamline detectors are not sensitive enough to  discriminate among the lighter particles in the beam: electrons, muons and pions fall under the same mass hypothesis. Thus, we need to assess the contamination from beamline particles other than pions in the event selections used for the pion cross section analysis and correct for this background. The first step of this process is assessing the percentage of electrons and muons in the $\pi/\mu/e$ beamline candidates via the G4Beamline MC, as we deem the percentage of kaons and antiprotons negligible after the mass selection (< 1 \textperthousand). 
Since the beamline composition is a function of the magnet settings, we simulate separately events for magnet current of -60A and -100A. 
Figure \ref{fig:BeamComposition} shows the momentum predictions from G4Beamline overlaid with data for the 60A runs (left) and for the 100A runs (right). The predictions for electrons, muons and pions have been staggered and their sum is area normalized to data. Albeit not perfect, these plots show a reasonable agreement in momentum shapes between data and MC.  We attribute the difference in shape (longer tail in data) to a two approximations performed in the MC. Firstly, G4Beamline lacks the simulation of the WC efficiency which is momentum dependent and leads to enhance the number events in the center of the momentum distribution. Secondly, G4Beamline stop tracking pions and their products if they decay after WC1; in data, pion decays in flight can still create a tigger if the produced muon travels thought the beamline detectors. In the pion cross section analysis, these differences between data and the G4Beamline prediction are accounted for as a systematic uncertainty related to the beam composition (see Section \ref{ch:BKGsubXSPuppa}).
%We attribute  the difference in shape to the lack of simulation of the WC efficiency in the MC which is momentum dependent and leads to enhance the number events in the center of the momentum distribution.

@@

\subsection{Data Driven MC}\label{sec:DDMC}
The Data Driven single particle Monte Carlo (DDMC) is a single particle gun which simulates the particle transport from WC4 into the TPC leveraging on the beamline data information. The DDMC uses the data momentum and position at WC4 to derive the event generation: a general sketch of the DDMC workflow is shown in Figure \ref{fig:DDMCSketch}.

When producing a DDMC sample, beamline data from a particular running period and/or running condition are selected first. For example, data for the negative 60A runs and for the negative 100A runs inform the event generation stage of two different DDMC samples. Figure \ref{fig:DDMCQuantities}  schematically shows the data quantities of interest leveraged from data: the momentum ($P_x, P_y, P_z$) and position ($X, Y$) at WC4. For each data event, we obtain the  particle position ($X, Y$) at WC4 directly from the data measurement; we calculate the components of the momentum using the beamline measurement of the momentum magnitude in conjunction with the hits on WC3 and WC4 to determine the direction of the momentum vector, as described in section \ref{sec:MWPCfunc}. The momentum and position of the selected data events form a 5-dimensional series of tuples.  The DDMC event generator samples from the joint distribution of these five quantities using a 5-dimensional hit-or-miss sampling procedure. This sampling generates MC events  with the same momentum and position distributions as data, with the additional benefit of accounting for the correlations between the $P_x, P_y, P_z, X, Y$ variables.  As an example, the results of the DDMC generation compared to data for the kaon +100A sample are shown in figure \ref{fig:DDMCComparison} for the $P_z$, $X$ and $Y$ distributions; as expected, MC and data agree within the statistical uncertainty by construction. A LArSoft simulation module then launches single particle MC from z = -100 cm (the location of the WC4) using the generated events. The particles are free to decay and interact in their path from WC4 to the TPC according to the Geant4 simulation.

Using the DDMC technique ensures that the MC and data particles have very similar momentum, position and angular distributions at WC4 and allows us to use the MC sample in several occasions: to estimate the background contamination to the pion cross section (see Section \ref{ch:PionXSBkgSub}), to calibrate the energy loss upstream of the TPC (see Section \ref{ch:eloss}),  and to study the tracking and the calorimetric performance  in the LArTPC (sections \ref{sec:TrackingStudies} and \ref{ch:energyCal}). A small caveat is in order here: the DDMC is a single particle Monte Carlo, which means that the beam pile-up is not simulated.

@@

We use the beamline simulation and the DDMC simulation to estimate the background in the total hadronic pion cross section. Two categories of background exists for the negative pion cross section measurement: the background related to the pion alternative processes to hadronic interaction inside the LArTPC, discussed in Section \ref{ch:CaptureAndDecay} and the background related to the beamline contamination, discussed in Section \ref{ch:PionXSBkgSub2}.

\subsection{Background from Pion Capture and Decay}\label{ch:CaptureAndDecay}
Our goal is to measure the total hadronic cross section for negative pions in argon. Since pion capture can be classified as an electromagnetic process and pion decay is a weak process (as discussed in Section \ref{ch:signature}),  capture and decay represent alternative processes to the hadronic interaction, resulting therefore into a background for the cross section analysis. We present here a study of capture and decay in Monte Carlo and the solution we adopted to remove their occurrence in the data sample.

For this MC study, we use a sample of  MC pions generated according to the  $-6$0A beam profile with the DDMC (see Section \ref{sec:DDMC}). It is important to notice that capture occurs predominantly at rest, while decay may occur both in flight and at rest. Thus, we can highly mitigate capture and decay at rest by removing pions whose kinetic energy at the TPC front face is low enough to be completely released by ionization in the TPC, eventually bringing the pion to stop within the fiducial volume. This translates into a beam momentum selection, where we keep only events whose WC momentum is above a certain threshold. 
Figure \ref{fig:CaptureMom} shows the true momentum distribution for the primary pions\footnote{We use here the Geant4 denomination ``primary" to indicate that the pion considered does not undergo interactions modifying its energy before getting to the TPC. In fact, not every pion shot from wire chamber four will arrive to the TPC as primary,  some will decay or interact before the TPC.}  that arrive to the TPC (pink), that capture (green) or decay (blue) inside the TPC, on a linear scale (left) and on a log scale (right) vertical axis.

In order to choose the selection value for the wire chamber momentum, it is beneficial to estimate the fraction of MC capture and decay events that survive the momentum selection as a function of the momentum threshold, and compare it with the survival fraction for all the 60A events. This is done in figure \ref{fig:survRatio}. We define the survival ratio simply  as the number of events surviving the true momentum selection divided by the number of events of that category. We calculate the survival ratio separately for the three event categories explained above: total (pink), capture (green) and decay (blue).
Selecting pions with momentum greater than 420 MeV/c removes $\sim$99\% the capture events while maintaining about 80\% of the 60A data sample and almost the entire 100A sample. 
Figure \ref{fig:evtRatio} shows the ratio of events which end their life in capture (green) or decay (blue) over the total number of events as a as a function of the true MC momentum at WC4. This ratio is slightly dependent on the inelastic cross section implemented in Geant4, as we are able to register a pion capture (or decay) only if it did not interact inelastically in the TPC. We choose a momentum threshold of 420~MeV/c because the percentage of capture events drops below 1\% and the percentage of decays is never above 2\% for momenta greater than 420~MeV/c. After the momentum selection, we evaluate the contribution of capture and decay to be a negligibly small background to the cross section measurement compared to the background related to the beamline which we will address in the next section.

\begin{figure}[]
\centering
\includegraphics[width=15cm]{Chapter-7/Images/CDAsMomentumFunct.png}
\caption{MC momentum distribution at wire chamber 4 for every simulated pion arriving in the TPC (pink), ending its life in capture (green) or in decay (blue) in the TPC, linear vertical axis on the left, logarithmic on the right. }
\label{fig:CaptureMom}
\end{figure}

\begin{figure}[]
\centering
\begin{minipage}[t]{0.45\textwidth}
\centering
\includegraphics[width=7.5cm]{Chapter-7/Images/CDThreshold.png}
\caption{Survival ratio as a function of selection threshold on MC momentum at wire chamber four for for every simulated pion arriving in the TPC (pink), capture (green) or in decay (blue).   }
\label{fig:survRatio}
\end{minipage}\hfill
\begin{minipage}[t]{0.45\textwidth}
\centering
\includegraphics[width=7.5cm]{Chapter-7/Images/CDRatio.png}
\caption{Ratio between the capture (green) and decay (blue) events over the total number of events as a as a function of the MC momentum at wire chamber four.}
\label{fig:evtRatio}
\end{minipage}
\end{figure}

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Contributions from the Beamline Background}\label{ch:PionXSBkgSub2}
We define beamline background every TPC track matched to the WC track which is not a primary pion. Potentially, there are 4 different types of beamline background:
\begin{itemize}
\item[]1) electrons,
\item[]2) muons,
\item[]3) secondaries from pion interactions in beamline elements,
\item[]4) matched pile up events.
\end{itemize}

The first step to quantify the effect of the beamline background on the pion cross section is to estimate what percentage of events used in the cross section calculation is not a primary pion.  We start by noting that the last type of background, the ``matched pile up" events, is a negligible fraction, because of the definition of the WC2TPC match: we deem the probability of a single match with a halo particle in the absence of a beamline particle\footnote{ Events with multiple WC2TPC matches are always rejected.} negligibly small. %\textcolor{red}{SHOW VTX distribution in WC2TPC match}
As shown in Section \ref{ch:beamlineComposition}, we use G4Beamline to estimate the percentage of pions, muons and electrons at WC4, obtaining the composition shown in Table \ref{tab:beamline}. The next step is to simulate those pions, muons and electrons from WC4 to the TPC with the DDMC and evaluate their contribution to the cross section. To do so, we start by generating the same number of electrons, muons and pions with the DDMC and we apply the same selection chain applied in data (i.e. track multiplicity rejection, WC2TPC match and shower rejection) on the three samples. The number of events per particle species surviving this selection is shown on table \ref{tab:MCafterCutContaminants}. In order to reproduce the data beamline composition, we weight each event of a given particle species according to the estimated content for that species as found with the G4Beamline simulation. In case of 60A runs, for example, the weights are 0.688 for pions,  0.046 for muons  and 0.266 for electrons.

It should be noted that pions may  interact hadronically in the steel (cryostat wall) or in the non-instrumented argon upstream to the TPC front face while traveling the length of between WC4 and the TPC. Or, they could decay in flight between WC4 and the TPC. One of the interaction or decay products can leak into the TPC and be matched with the WC track, contributing to the pool of events used for the cross section calculation. We call this type of particles ``secondaries" from pion events, with a terminology inspired by Geant4.  We estimate the number of secondaries using the DDMC pion sample.  The percentage of secondaries is given by the number of matched WC2TPC tracks whose corresponding particle is not flagged as primary by Geant4.  The secondary particles to primary pion ratio is 4.9\% in the 60A sample and 4.3\% in the 100A sample.

@@

                                                  & MC $\pi^-$   & MC  $ \mu^-$ & MC  $e^-$ & MC  $\pi^-$ & MC  $\mu^-$ & MC  $e^-$  \\
\hline
&  &  &  & & &\\  
Total Initial events                     & 334500  & 334500 & 334500 &344500 &344500& 344500 \\
After Multiplicity Rejection        & 330668  & 333420 & 198065 &326576 &344208& 201380 \\
After WC2TPC Selection          & 218239  & 296333 & 91139  &230418 &300228& 98834  \\
Evts After Shower Rejection     & 208063  & 288914 &  20293 &219882 &293585& 17780  \\
&  &  &  & & &\\  
  \hline
&  &  &  & & &\\  
Selection Survival Rate           &62.2\% & 86.4\% & 6.1\% & 63.8\%& 85.2\%& 5.2\%\\
Beam Composition  @WC4      &  68.8\%   &  4.6 \%  & 26.6 \%    & 87.4 \% & 3.7 \%  & 8.9 \% \\ %
Expected Composition in XS sample &  88.5\%   & 8.2\%   & 3.3 \%   & 94.0\%	& 5.3\% & 0.7\%\\
                                                  &                      &                       &                   &                       &                        &\\  
\hline
\end{tabular}
\caption{MC selection flow per particle species.}
\label{tab:MCafterCutContaminants}
\end{table}

@@

\begin{figure}[p]
\centering
\includegraphics[width=\textwidth]{Chapter-5/Images/Background100A.pdf}
\caption{Left: staggered contributions to the interacting kinetic energy distribution for electron (grey), muons (yellow) and pion (blue) in the 100A simulation sample. Right: staggered contributions to the incident kinetic energy distribution for electron (grey), muons (yellow) and pion (blue) in the 100A simulation sample.  }
\label{fig:stag100A}
\end{figure}

We evaluate the beamline background contribution to the cross section by producing the interacting and incident histograms for the signal and background events surviving the selection, staggering the contributions for each particle species, as shown in Figure  \ref{fig:stag60A} for the -60A case and  in Figure \ref{fig:stag100A} for the -100A case. From those histograms, we are able to evaluate the contribution of  pions and  beamline backgrounds to each bin of the interacting and incident histograms separately and obtain the relative pion content. The relative pion content in each bin for the interacting and incident histograms represents the correction applied to data. We take here the interacting histogram as example, noting that the derivation of the correction for the incident histogram is identical. The number of entries in each bin of the interacting histogram (Figure \ref{fig:stag60A} left) is  $N^{\text{TOT}}_{\text{Int}} (E_{i})$, equal to the sum of the pions and beamline backgrounds flagged as interacting in that bin, namely

@@

The pion content is evaluated separately in the interacting and incident histograms. Their ratio determines a correction to the measured raw cross section. 
For example, the measured raw  cross section of a sample with enhanced muons content will tend to be lower than the raw cross section of a muon free sample. This is because most of the muons will cross the TPC without stopping, thus contributing almost exclusively to the incident histogram, forcing the pion content to be lower in the incident histogram than in the interacting; thus, the correction will tend to enhance the cross section. We present the estimation of $C^{\pi MC}_{\text{Int}} (E_{i})$ and $C^{\pi MC}_{\text{Inc}} (E_{i})$ and their associated systematic uncertainty in chapter discussing the pion cross section measurement (Figure \ref{fig:CorrectionsBeam}).

@@

\section{Tracking Studies}\label{sec:TrackingStudies}
The tracking of hadrons in the TPC determines both the beamline to TPC handshake and the identification of the interaction point within the TPC. Thus, it plays a fundamental role in the cross section measurements. We performed several studies geared towards the optimization of the LArSoft software package for tracking in the TPC. In particular, we studied a suitable set of parameters for the WC2TPC match and we optimized the clustering algorithm to maximize the efficiency of finding the interaction point on MC. Given the technical nature of these studies, we report them in Appendix \ref{ch:AppendixTrack}. 
We only report here the evaluation of  the angular resolution of the tracking algorithm in data and MC, due to its important implications on the physics measurement.

\subsection{Angular Resolution}\label{sec:angleRes}
Scope of this study is to understand and compare the tracking performances and the angular resolution of the TPC tracking on data and MC. 
We use the angular resolution of the tracking to determine  the value of smallest angle that the tracking algorithms allow to reconstruct with a non-zero efficiency, effectively determining a selection on the distribution of the scattering angle of hadronic interaction entering the cross section measurement. %\textcolor{red}{This study is performed on the pion sample, but its results are extrapolated to the kaon case.}

We start by selecting all the WC2TPC matched tracks used for the cross section analysis.  These tracks can contain from a minimum of 3 3D-space points to a maximum of 240  3D-space points.  We fit a straight line to all the 3D-space points associated with the track. 
For each track we calculate the average distance between each point in space and the fit line as follows 
\begin{equation} 
\bar d = \frac{\sum^N_i d_i}{N},
\end{equation} 
where $N$ is the number of 3D-space points of the track and $d_i$ is the distance of the $i$-th space point to the line fit. Several tests to compare the goodness of fit between data and MC have been considered. We decided to use $\bar d$ for its straightforward interpretation. The $\bar d$ distribution for data and MC is shown in Figure \ref{fig:Chi2AllPts} for pions and in Figure \ref{fig:Chi2AllPtsK} for kaons and shows a relatively good agreement between data and MC.
